---
title: 'Project 2: Modeling and Evaluation'
subtitle: "CSE6242 - Data and Visual Analytics - Summer 2018\n\nDue: Sunday, July 24, 2018 at 11:59 PM UTC-12:00 on T-Square"
output:
  pdf_document: default
  html_document:
    code_folding: none
    theme: default
  html_notebook:
    code_folding: none
    theme: default
---

# Data

We will use the same dataset as Project 1: [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged).

# Objective

Your goal in this project is to build a linear regression model that can predict the `Gross` revenue earned by a movie based on other variables. You may use R packages to fit and evaluate a regression model (no need to implement regression yourself). Please stick to linear regression, however.

# Instructions

You should be familiar with using an [RMarkdown](http://rmarkdown.rstudio.com) Notebook by now. Remember that you have to open it in RStudio, and you can run code chunks by pressing *Cmd+Shift+Enter*.

Please complete the tasks below and submit this R Markdown file (as **pr2.Rmd**) containing all completed code chunks and written responses, and a PDF export of it (as **pr2.pdf**) which should include the outputs and plots as well.

_Note that **Setup** and **Data Preprocessing** steps do not carry any points, however, they need to be completed as instructed in order to get meaningful results._

# Setup

Same as Project 1, load the dataset into memory:

```{r}
set.seed(2468)
load('movies_merged')
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

## Load R packages

Load any R packages that you will need to use. You can come back to this chunk, edit it and re-run to load any additional packages later.

```{r}
library(ggplot2)
library(GGally)
library(ggplot2)
library(GGally)
library(assertr)
library(splitstackshape)
library(lubridate)
library(reshape2)
library(plyr)
```

If you are using any non-standard packages (ones that have not been discussed in class or explicitly allowed for this project), please mention them below. Include any special instructions if they cannot be installed using the regular `install.packages('<pkg name>')` command.

**Non-standard packages used**: None

# Data Preprocessing

Before we start building models, we should clean up the dataset and perform any preprocessing steps that may be necessary. Some of these steps can be copied in from your Project 1 solution. It may be helpful to print the dimensions of the resulting dataframe at each step.

## 1. Remove non-movie rows

```{r}
# TODO: Remove all rows from df that do not correspond to movies
df <- df[which(df$Type == 'movie'),]
dim(df)
```

## 2. Drop rows with missing `Gross` value

Since our goal is to model `Gross` revenue against other variables, rows that have missing `Gross` values are not useful to us.

```{r}
# TODO: Remove rows with missing Gross value
df <- df[!is.na(df$Gross),]
dim(df)
```

## 3. Exclude movies released prior to 2000

Inflation and other global financial factors may affect the revenue earned by movies during certain periods of time. Taking that into account is out of scope for this project, so let's exclude all movies that were released prior to the year 2000 (you may use `Released`, `Date` or `Year` for this purpose).

```{r}
# TODO: Exclude movies released prior to 2000
# Extract Year from 'Released_Year', if Released_Year is blank fill it up with value from 'Date' column
df$Released_year <- year(ymd(df$Released))
df$Released_year[is.na(df$Released_year)] <- df$Year[is.na(df$Released_year)]
df <- df[which(df$Released_year >= 2000), ]
#df$month <-  month(ymd(df$Released))
dim(df)
```

## 4. Eliminate mismatched rows

_Note: You may compare the `Released` column (string representation of release date) with either `Year` or `Date` (numeric representation of the year) to find mismatches. The goal is to avoid removing more than 10% of the rows._

```{r}
# TODO: Remove mismatched rows
# df1 <- df[(df$Year == df$Released_year),]
df$diff_year <- abs(df$Released_year - df$Year)
df <- df[which(df$diff_year < 2), ] #removing rows when there is more than one year diff between Year and Release Year.
df$diff_year <- NULL
dim(df)
```

## 5. Drop `Domestic_Gross` column

`Domestic_Gross` is basically the amount of revenue a movie earned within the US. Understandably, it is very highly correlated with `Gross` and is in fact equal to it for movies that were not released globally. Hence, it should be removed for modeling purposes.

```{r}
# TODO: Exclude the `Domestic_Gross` column
df$Domestic_Gross <- NULL
dim(df)
```

## 6. Process `Runtime` column

```{r}
# TODO: Replace df$Runtime with a numeric column containing the runtime in minutes
parse_runtime <- function(runtime){
  runtime <- strsplit(runtime, ' ')[[1]]
  if((length(runtime) == 2) && (runtime[2] == 'h')){
    return(as.integer(runtime[1])*60)
  }else if((length(runtime) == 2) && (runtime[2] == 'min')){
    return(as.integer(runtime[1]))
  }else if(length(runtime) == 4){
    hr_to_min = as.integer(runtime[1])*60
    return(hr_to_min + as.integer(runtime[3]))
  }else return(NA)
}
df$Runtime <- sapply(df$Runtime, parse_runtime)
```

Perform any additional preprocessing steps that you find necessary, such as dealing with missing values or highly correlated columns (feel free to add more code chunks, markdown blocks and plots here as necessary).

## 7. Additional preprocessing - NA values

```{r}
# TODO(optional): Additional preprocessing on Numeric Columns

df_numeric <- df[,sapply(df, function(x) is.numeric(x))]
df_numeric <- cbind(df$Title, df_numeric)
colnames(df_numeric)[1] <- "Title"
df_numeric$Year <- NULL
df_numeric$Date <- NULL
df_numeric$diff_year <- NULL
df_numeric$Metascore <- as.numeric(as.character(df$Metascore))

# check NA values in all column
print(sapply(df_numeric, function(x) sum(is.na(x))))
# Remove rows where Runtime is not available
df_numeric <- df_numeric[!is.na(df_numeric$Runtime), ] # 37 rows removed

#df_NA <- na.omit(df)
#dim(df_NA)

# for real values features, replace NA with "average""
# tomatoRating and tomatoUserRating (5, 10 cols)
for(i in 2:ncol(df_numeric)){
  for (j in c(5,10)){
    df_numeric[is.na(df_numeric[,i]), j] <- mean(df_numeric[,i], na.rm = TRUE)
  }
}

# For the rest of NA values which are count replace them with "median""
for(i in 2:ncol(df_numeric)){
  df_numeric[is.na(df_numeric[,i]), i] <- median(df_numeric[,i], na.rm = TRUE)
}

# Re-check if no NA values in all column
print(sapply(df_numeric, function(x) sum(is.na(x))))

df <- merge(df, df_numeric['Title'], by = 'Title')
```

_**Note**: Do NOT convert categorical variables (like `Genre`) into binary columns yet. You will do that later as part of a model improvement task._

## Final preprocessed dataset

Report the dimensions of the preprocessed dataset you will be using for modeling and evaluation, and print all the final column names. (Again, `Domestic_Gross` should not be in this list!)

```{r}
# TODO: Print the dimensions of the final preprocessed dataset and column names
print(cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file=""))
print(colnames(df))
```


# Evaluation Strategy

In each of the tasks described in the next section, you will build a regression model. In order to compare their performance, you will compute the training and test Root Mean Squared Error (RMSE) at different training set sizes.

First, randomly sample 10-20% of the preprocessed dataset and keep that aside as the **test set**. Do not use these rows for training! The remainder of the preprocessed dataset is your **training data**.

Now use the following evaluation procedure for each model:

- Choose a suitable sequence of training set sizes, e.g. 10%, 20%, 30%, ..., 100% (10-20 different sizes should suffice). For each size, sample that many inputs from the training data, train your model, and compute the resulting training and test RMSE.
- Repeat your training and evaluation at least 10 times at each training set size, and average the RMSE results for stability.
- Generate a graph of the averaged train and test RMSE values as a function of the train set size (%), with optional error bars.

You can define a helper function that applies this procedure to a given set of features and reuse it.

# Tasks

Each of the following tasks is worth 20 points, for a total of 100 points for this project. Remember to build each model as specified, evaluate it using the strategy outlined above, and plot the training and test errors by training set size (%).

## 1. Numeric variables

Use Linear Regression to predict `Gross` based on available _numeric_ variables. You can choose to include all or a subset of them.

```{r}
# TODO: Build & evaluate model 1 (numeric variables only)
df_num <- df_numeric #for feature transformations
#remove(df_numeric)

RMSE <- function(data, train_size){
  average_rmse_train <- c()
  average_rmse_test <- c()
  iter = 0
  # shuffle and split train and test datasets - 90/10
  n_train <- ceiling(nrow(data)*0.9) 
  shuffle <- sample(nrow(data))
  train_actual <- data[shuffle[1:n_train], ]
  test_df <- data[shuffle[(n_train+1):nrow(data)],]
  
  while(iter < 20){
    n_train_pct <- ceiling(nrow(train_actual)*train_size)
    shuffle_train <- sample(nrow(train_actual))
    train_df <- train_actual[shuffle_train[1:n_train_pct], ]   
    #cat(rankMatrix(train_df), "\n")
    #cat(rankMatrix(test_df), "\n")
    lm.out <- lm(Gross~., data = train_df)
    pred_train <- suppressWarnings(predict(lm.out, train_df))
    pred_test <- suppressWarnings(predict(lm.out, test_df))

    #pred_train <- predict(lm.out, train_df)
    #pred_test <- predict(lm.out, test_df)

    average_rmse_train <- append(average_rmse_train, sqrt(mean((train_df$Gross - pred_train)^2))) 
    average_rmse_test <- append(average_rmse_test, sqrt(mean((test_df$Gross - pred_test)^2)))
    
    #average_rmse_train <- append(average_rmse_train, RMSE(train_df$Gross, pred_train)) 
    #average_rmse_test <- append(average_rmse_test, RMSE(test_df$Gross, pred_test))

    iter = iter + 1
  }
  rmse_train = mean(average_rmse_train)
  rmse_test = mean(average_rmse_test)
  #print(rmse_test)
  return(c(rmse_train, rmse_test))
}

# Removing Title from numeric dataset
df_numeric_train <- subset(df_numeric, select = -Title)

# collecting RMSE stats
training_size <- seq(0.1, 1, length.out = 15)
training_size_stats <- data.frame(matrix(nrow = 15, ncol =3))
colnames(training_size_stats) <- c("training_size", "train_rmse", "test_rmse")

for(i in 1:length(training_size)){
  resutls_0_1 <- RMSE(df_numeric_train, training_size[i])
  training_size_stats[i,] <- c(training_size[i], resutls_0_1[1], resutls_0_1[2])
}

#lm_new_num <- lm(Gross~., data = test)
#summary(lm_new_num)

rmse_task1 <- melt(training_size_stats[,c('training_size', "train_rmse", "test_rmse")], id = "training_size")
colnames(rmse_task1) <- c("training_size", "variable", "RMSE")

ggplot(rmse_task1, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
  geom_line(linetype = "dashed") + 
  geom_point() + scale_y_log10() +
ggtitle("Numeric Feature Modelling")

#print(colnames(df_numeric_train))
print(paste("Best mean test RMSE value observed for numeric features:", min(training_size_stats$test_rmse), sep = " "))
```

**Q**: List the numeric variables you used.

**A**:  [1] "Runtime"           "imdbRating"        "imdbVotes"         "tomatoMeter"       "tomatoRating"      "tomatoReviews"    
        [7] "tomatoFresh"       "tomatoRotten"      "tomatoUserMeter"   "tomatoUserRating"  "tomatoUserReviews" "Budget"           
        [13]"Released_year"     "Metascore"         "Gross" (Target)             



**Q**: What is the best mean test RMSE value you observed, and at what training set size?
**A**: Best mean test RMSE value observed for numeric features: 83858702.8399003, training_size is 87.15%


## 2. Feature transformations

Try to improve the prediction quality from **Task 1** as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g. `is_budget_greater_than_3M`).

```{r message = FALSE, warnings = FALSE}
# TODO: Build & evaluate model 2 (transformed numeric variables only)

### Power transformations should be before one-hot and binning
num_col <- colnames(df_numeric_train)
num_col <- num_col[! num_col %in% 'Gross']
power_df <- data.frame(matrix(NA, nrow = nrow(df_numeric_train), ncol = 0))

for(j in 1:length(num_col)){
    #power_df[num_col[j]] <- df_numeric_train[,num_col[j]]
    log_new_col <- log10(df_numeric_train[,num_col[j]] + 0.01) #to avoid log10(0)
    power_df[, paste0('log_', num_col[j])] <- log_new_col
    sqr_new_col <- (df_numeric_train[,num_col[j]])^2
    power_df[, paste0('square_', num_col[j])] <- sqr_new_col
    cub_new_col <- (df_numeric_train[,num_col[j]])^3
    power_df[, paste0('cube_', num_col[j])] <- cub_new_col
}

new_power_cols <- colnames(power_df)

# Replace NA values with median data, shoudldnt be any
for(i in 1:ncol(power_df)){
  power_df[is.na(power_df[,i]), i] <- median(power_df[,i], na.rm = TRUE)
}

#normalize all the columns
power_df <- data.frame(sapply(power_df, function(x) scale(x, center = TRUE))) 
power_df['Gross'] <- df_numeric_train$Gross

# Select the significant transformations
lm_new_num <- lm(Gross~., data = power_df)
summary(lm_new_num)

power_df$Gross <- NULL
combined_df <- cbind(df_numeric, power_df)
#unique_cols <- unique(colnames(combined_df)


feature_signif <- c('square_tomatoReviews',
'cube_tomatoReviews',
'square_tomatoRotten',
'cube_tomatoUserMeter',
'square_tomatoUserReviews',
'cube_tomatoUserReviews',
'square_imdbVotes',
'log_imdbVotes',
'cube_tomatoUserMeter',
'cube_Budget',
'log_Released_year',
'square_Released_year')

combined_cols_wanted <- c(colnames(df_numeric), feature_signif)
combined_df_wanted <- combined_df[,combined_cols_wanted]
combined_df_wanted$cube_tomatoUserMeter.1 <- NULL

#combined_df_wanted <- combined_df_wanted_withTitle
combined_df_wanted_withTitle <- combined_df_wanted #saved for later
#combined_df_wanted$Title <- NULL

#lm_new_num <- lm(Gross~., data = combined_df_wanted)
#summary(lm_new_num)
#combined_df_wanted <- combined_df #not removing any features


# budget quantiles
budget_quantiles <- quantile(combined_df_wanted$Budget, probs = seq(0, 1, 0.2), na.rm = TRUE)
budgetQuantiles <- function(budget) {
  cut(budget, breaks = budget_quantiles, labels = c("Ultra-Low-Budget", "Low-Budget", "Medium-Budget", "Large-Budget", "Ultra-Large-Budget"))
}
combined_df_wanted$budget_quant <- sapply(combined_df_wanted$Budget, budgetQuantiles)
combined_df_wanted <- cSplit_e(combined_df_wanted, "budget_quant", ",", type = "character", fill = 0) #split Genre column in to 
combined_df_wanted$`budget_quant_Ultra-Low-Budget` <- NULL #to remove linearly dependent vars)

# runtime quantiles
#runtime_quantiles <- quantile(combined_df_wanted$Runtime, probs = seq(0, 1, 0.33), na.rm = TRUE)
#runtimeQuantiles <- function(runtime) {
#  cut(runtime, breaks = runtime_quantiles, labels = c( "Short", "Medium", "Long"))
#}
#combined_df_wanted$runtime_quant <- sapply(combined_df_wanted$Runtime, runtimeQuantiles)
#combined_df_wanted <- cSplit_e(combined_df_wanted, "runtime_quant", ",", type = "character", fill = 0) #split Genre column in to 
#combined_df_wanted$runtime_quant_Short <- NULL #to remove linearly dependent vars)

### One hot encoding and binning
# imdb quantiles
imdb_quantiles <- quantile(combined_df_wanted$imdbRating, probs = seq(0, 1, 0.33), na.rm = TRUE)
imdbQuantiles <- function(imdbrating) {
  cut(imdbrating, breaks = imdb_quantiles, labels = c( "bad", "average", "good"))
}
combined_df_wanted$imdbrating_Q <- sapply(combined_df_wanted$imdbRating, imdbQuantiles)
combined_df_wanted <- cSplit_e(combined_df_wanted, "imdbrating_Q", ",", type = "character", fill = 0) #split Genre column in to 
combined_df_wanted$imdbrating_Q_good <- NULL #to remove linearly dependent vars)

# tomatoRating quantiles
tomato_quantiles <- quantile(combined_df_wanted$tomatoRating, probs = seq(0, 1, 0.33), na.rm = TRUE)
tomatoQuantiles <- function(tomatorating) {
  cut(tomatorating, breaks = tomato_quantiles, labels = c( "bad", "average", "good"))
}

combined_df_wanted$tomatorating_Q <- sapply(combined_df_wanted$tomatoRating, tomatoQuantiles)
combined_df_wanted <- cSplit_e(combined_df_wanted, "tomatorating_Q", ",", type = "character", fill = 0) #split Genre column in to 
combined_df_wanted$tomatorating_Q_good <- NULL #to remove linearly dependent vars)

#runtime bins
runtime<-combined_df_wanted$Runtime
combined_df_wanted$Runtime_Bin<-cut(runtime,seq(min(runtime)-1,max(runtime)+1,(max(runtime) - min(runtime)+2)/6),right=FALSE,labels=c(1:6))
combined_df_wanted$Runtime_Bin <- as.numeric(as.factor(combined_df_wanted$Runtime_Bin))

metascore <- combined_df_wanted$Metascore
combined_df_wanted$metascore_bin<-cut(metascore,seq(min(metascore)-1,max(metascore)+1,(max(metascore) - min(metascore)+2)/6),right=FALSE,labels=c(1:6))
combined_df_wanted$metascore_bin <- as.numeric(as.factor(combined_df_wanted$metascore_bin))

imdbvotes <- combined_df_wanted$imdbVotes
combined_df_wanted$imdbvote_bin<-cut(imdbvotes,seq(min(imdbvotes)-1,max(imdbvotes)+1,(max(imdbvotes) - min(imdbvotes)+2)/15),right=FALSE,labels=c(1:15))
combined_df_wanted$imdbvote_bin <- as.numeric(as.factor(combined_df_wanted$imdbvote_bin))

# removing na in categoriacal like tomato_Q
combined_df_wanted <- na.omit(combined_df_wanted) #Has Title
# select only numerical, shouldnt be any other than Title
temp <- combined_df_wanted[,sapply(combined_df_wanted, function(x) is.numeric(x))]

# 'training_size_stats_tranform7' - with only wanted cols and no quantiles - looks better
training_size <- seq(0.1, 1, length.out = 15)
training_size_stats_tranform7 <- data.frame(matrix(nrow = 10, ncol =3))
colnames(training_size_stats_tranform7) <- c("training_size", "train_rmse", "test_rmse")

for(i in 1:length(training_size)){
  resutls_transform <- RMSE(temp, training_size[i])
  training_size_stats_tranform7[i,] <- c(training_size[i], resutls_transform[1], resutls_transform[2])
}

rmse_task2 <- melt(training_size_stats_tranform7[,c('training_size', "train_rmse", "test_rmse")], id = "training_size")
colnames(rmse_task2) <- c("training_size", "variable", "RMSE")

ggplot(rmse_task2, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
  geom_line(linetype = "dashed") + 
  geom_point() + scale_y_log10() +
ggtitle("Transformed Numeric Feature Modelling")


print(paste("Best mean test RMSE value observed for numeric features:", min(training_size_stats_tranform7$test_rmse), sep = " "))
```

**Q**: Explain which transformations you used and why you chose them.

**A**: 1. Binning: Binning's really only a good idea when you'd expect a discontinuity in the response at the cut-points. Runtime although a real valued feature there are discontinuities i.e Runtime generally fall in to few buckets like short, medium, long, very long e.t.c

      2. One-hot Encoding: In the EDA from Project 1, we have observed distinct distributions of Budget, hence I created 5 categories of budget i.e. Low Budget, High Budget, Large Budget e.t.c as one-hot encoded variables

      3. Power Transforming: I have tested log, squared and power transformations on all the numeric variables. By looking at plots of features like Budget and looking at the significane of transformed features, I have filtered them to a set of 11 new features. 

**Q**: How did the RMSE change compared to Task 1?

**A**: R^2 increased from 0.72 to 0.76 when added these transformed features. Overall the test RMSE has improved compared to Pure numeric feature test RMSE at training size greater than 40%. Best mean test RMSE value observed for numeric features: 71969667.8478177 at 42% and second best test RMSE 74596218 at 90% training size.


## 3. Non-numeric variables

Write code that converts genre, actors, directors, and other categorical variables to columns that can be used for regression (e.g. binary columns as you did in Project 1). Also process variables such as awards into more useful columns (again, like you did in Project 1). Now use these converted columns only to build your next model.

```{r}
# TODO: Build & evaluate model 3 (converted non-numeric variables only)

#### Convert top 12 Genre to binary features, 12 Genre's account for more than 85% data
df1 <- df #delete later

#df_categ <- data.frame(matrix(NA, nrow = nrow(df_num), ncol = 0))

df_categ <- df[, sapply(df, function(x) !is.numeric(x))] #replace all df with df_char for categorical features
df_categ$Released_year <- df$Released_year

genre_df <- df_categ[,c('Title', 'Genre', 'Released_year')]
genre_df <- cSplit_e(genre_df, "Genre", ",", type = "character", fill = 0) #split Genre column in to 
genre_df$`Genre_N/A` <- NULL

# add month from date columns
df_categ$month <-  month(ymd(df$Released))
df_categ$Released <- NULL

# tomato Image
# if the positive reviews are less than 60%, the film is considered "rotten". ... The "Certified Fresh" seal is reserved for movies that satisfy two criteria: a "Tomatometer" of 75% or better and at least 40 reviews (for limited release movies, otherwise 80) from "Tomatometer" critics (including 5 Top Critics).

df_categ <- cSplit_e(df_categ, "Rated", ",", type = "character", fill = 0) #split Genre column in to 
df_categ$`Rated_N/A` <- NULL
df_categ$`Rated_NOT RATED` <- NULL

# movie rating
df_categ <- cSplit_e(df_categ, "tomatoImage", ",", type = "character", fill = 0) #split Genre column in to 
df_categ$`tomatoImage_N/A` <- NULL

most_cmon_10 <-  as.data.frame(sort(colSums(genre_df[,4:ncol(genre_df)]), decreasing = TRUE)[1:12], row.names = NULL)
most_cmon_10$Genre <- rownames(most_cmon_10)
colnames(most_cmon_10) <- c("Count", "Genre")
#most_cmon_10$Genre <- gsub("Genre_", "", most_cmon_10$Genre)
#colnames(genre_df) <- gsub("Genre_", "", colnames(genre_df))
req_cols = append(c('Title', 'Released_year') , most_cmon_10$Genre )
genre_df <- genre_df[, req_cols] #only keep top 10 most frequent Genres
df_categ$Genre <- NULL
df_categ <- merge(df_categ, genre_df, by = c('Title', 'Released_year'))

#### Convert Writer feature to no.of writers for features
#remove uncesseray characters from Writer feature
#df2 <- df_categ #delete later
#df_categ$Writer <- gsub(" \\([a-z]*\\)","", df_categ$Writer)

#writer_df <- df_categ[,c('Title', 'Writer', 'Released_year')]
#writer_df <- cSplit_e(writer_df, "Writer", ",", type = "character", fill = 0) #split Writer column in to 
#writer_df$`Writer_N/A` <- NULL

#df_categ$Writer <- NULL
#df_categ$Writer <-  as.data.frame(rowSums(writer_df[,4:ncol(writer_df)]))

#### Convert Production to binary feature
df3 <- df_categ #delete later

# Sony has different forms like Sony Pictures, Sony Columbia Pictures, Sony Entertainment
prod_df <- df[,c('Title', 'Production', 'Released_year')]
prod_df$Production <- tolower(prod_df$Production)
prod_df$Production <- gsub(".*sony.*", "sony pictures", prod_df$Production)
prod_df$Production <- gsub(".*warner.*", "warner bros", prod_df$Production)
prod_df$Production <- gsub(".*weinsten.*", "weinsten company", prod_df$Production)
prod_df$Production <- gsub(".*universal.*", "universal pictures", prod_df$Production)
prod_df$Production <- gsub(".*twentieth.*", "20th century fox", prod_df$Production)
prod_df$Production <- gsub(".*20th century.*", "20th century fox", prod_df$Production)

prod_df <- cSplit_e(prod_df, "Production", ",", type = "character", fill = 0) #split Genre column in to 
prod_df$`Production_n/a` <- NULL
prod_count <-  as.data.frame(sort(colSums(prod_df[,4:ncol(prod_df)]), decreasing = TRUE)[1:12], row.names = NULL)
prod_count$production <- rownames(prod_count)
colnames(prod_count) <- c("Count", "production")
#most_cmon_10$Genre <- gsub("Genre_", "", most_cmon_10$Genre)
#colnames(genre_df) <- gsub("Genre_", "", colnames(genre_df))
req_cols = append(c('Title', 'Released_year') , prod_count$production )
prod_df <- prod_df[, req_cols] #only keep top 11 movie production houses
df_categ$Production <- NULL
df_categ <- merge(df_categ, prod_df, by = c('Title', 'Released_year'))
df_categ$`Production_n/a` <- NULL

# Creating Wins and Nominations feature from Awards feature
awards_df <- df_categ
#awards_df <- awards_df[which(awards_df$Awards != 'N/A'), ]
# awards_df <- df[sample(nrow(awards_df), 5000), ] #sampling 5000 to reduce runtime for submission
parse_awards <- function(movie_db){
  
  regex_awards <- function(awards) {
    temp_x <- regmatches(awards, gregexpr("+[0-9]+", awards))[[1]]
    temp_x <- as.numeric(unlist(temp_x))
  }
  movie_db$numeric_awards <- sapply(movie_db$Awards, regex_awards)
  #function(awards) as.numeric(unlist(regmatches(awards, gregexpr("+[0-9]+", awards))[[1]]) )
  movie_db$Wins = 0
  movie_db$Nominations = 0
  Count_InvalidAwards <- 0 #
  cv <- nrow(movie_db)
  for(i in 1:cv){
    if( (length(movie_db[i,]$numeric_awards[[1]])!=3) & (length(movie_db[i,]$numeric_awards[[1]])!=2) & (length(movie_db[i,]$numeric_awards[[1]])!=0)){
      if(length(grep("win",movie_db[i,]$Awards))>0){
        movie_db[i,]$Wins = as.numeric(movie_db[i,]$numeric_awards[[1]][1])          
      }
      if(length(grep("nomination",movie_db[i,]$Awards))>0){
        movie_db[i,]$Nominations = as.numeric(movie_db[i,]$numeric_awards[[1]][1])   
      }
    }
    if(length(movie_db[i,]$numeric_awards[[1]])==2){
      if((length(grep("win",movie_db[i,]$Awards))>0) || (length(grep("won",movie_db[i,]$Awards))>0)|| (length(grep("Won",movie_db[i,]$Awards))>0)){
        movie_db[i,]$Wins = as.numeric(movie_db[i,]$numeric_awards[[1]][1])          
      }
      if(length(grep("nomination",movie_db[i,]$Awards))>0){
        movie_db[i,]$Nominations = as.numeric(movie_db[i,]$numeric_awards[[1]][2])   
      }
    }
    if(length(movie_db[i,]$numeric_awards[[1]])==3){
      if((length(grep("win",movie_db[i,]$Awards))>0) || (length(grep("won",movie_db[i,]$Awards))>0)|| (length(grep("Won",movie_db[i,]$Awards))>0)){
        movie_db[i,]$Wins = as.numeric(movie_db[i,]$numeric_awards[[1]][2])          
      }
      if(length(grep("nomination",movie_db[i,]$Awards))>0){
        movie_db[i,]$Nominations = as.numeric(movie_db[i,]$numeric_awards[[1]][3])   
      }
    }
    if(length(movie_db[i,]$numeric_awards[[1]])==0){
      movie_db[i,]$Wins = 0
      movie_db[i,]$Nominations = 0
      Count_InvalidAwards = Count_InvalidAwards + 1
    }
    
    if(length(unlist(movie_db[i,]$Wins))==1)
      movie_db[i,]$Wins = as.numeric(unlist(movie_db[i,]$Wins))
    else
      print(paste(movie_db[i,]$Awards, movie_db[i,]$Wins, sep = " ==> " ))
    if(length(unlist(movie_db[i,]$Nominations))==1)
      movie_db[i,]$Nominations = as.numeric(unlist(movie_db[i,]$Nominations)[[1]])
    else
      print(paste(movie_db[i,]$Nominations, movie_db[i,]$Nominations, sep = " ==> " ))
  }
  print("Invalid Number of Rows w/o excluding 'N/A' in Awards")
  print(Count_InvalidAwards)
  return(movie_db)
}
df_categ <- parse_awards(awards_df)
remove(awards_df)
df_categ$numeric_awards <- NULL

# Add Gross to df_categ for model training
df_categ <- merge(df_categ, df[,c('Title', 'Released_year', 'Gross')], by = c('Title', 'Released_year'))
df_categ$Released_year <- NULL

df_categ <- na.omit(df_categ) # used for Task 5

num_categ_cols <- sapply(df_categ, function(x) is.numeric(x))

df_categ_train <- df_categ[, num_categ_cols]


## model for categorical features
training_size <- seq(0.1, 1, length.out = 20)
training_size_stats_categ <- data.frame(matrix(nrow = 20, ncol =3))
colnames(training_size_stats_categ) <- c("training_size", "train_rmse", "test_rmse")

for(i in 1:length(training_size)){
  resutls_categ <- RMSE(df_categ_train, training_size[i])
  training_size_stats_categ[i,] <- c(training_size[i], resutls_categ[1], resutls_categ[2])
}

rmse_task3 <- melt(training_size_stats_categ[,c('training_size', "train_rmse", "test_rmse")], id = "training_size")
colnames(rmse_task2) <- c("training_size", "variable", "RMSE")

ggplot(rmse_task2, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
  geom_line(linetype = "dashed") + 
  geom_point() + scale_y_log10() +
ggtitle("Transformed Numeric Feature Modelling")


print(paste("Best mean test RMSE value observed for numeric features:", min(training_size_stats_categ$test_rmse), sep = " "))

```

**Q**: Explain which categorical variables you used, and how you encoded them into features.

**A**: 


**Q**: What is the best mean test RMSE value you observed, and at what training set size? How does this compare with Task 2?

**A**: 1. Best mean test RMSE value observed for numeric features: 120608544.579493 at 67% training size. Since we are only using categorical features and excluded very significant numerical features like budget, Categorical model has high RMSE compared to test RMSE from Task 2.

      2. R-Squared value is approx. .45 compared ~.76 from Task 2.

## 4. Numeric and categorical variables

Try to improve the prediction quality as much as possible by using both numeric and non-numeric variables from **Tasks 2 & 3**.

```{r}
# TODO: Build & evaluate model 4 (numeric & converted non-numeric variables)
#merge df_num and df_categ to df_all

#combined_df1 <- cbind(df$Title, combined_df)
#colnames(combined_df1) <- c("Title", colnames(combined_df))

combined_df1 <- cbind(df$Title, combined_df_wanted)
colnames(combined_df1) <- c("Title", colnames(combined_df_wanted))

combined_num_categ_df = merge(combined_df1, df_categ, by = "Title")

combined_num_categ_df = na.omit(combined_num_categ_df)

combined_num_categ_df = rename(combined_num_categ_df, c("Gross.x" = "Gross"))
combined_num_categ_df$Gross.y <- NULL
#combined_num_categ_df = rename(combined_num_categ_df, c("Writer" = "Writer"))
#colnames(combined_num_categ_df) <- colnames(combined_num_categ_df)

#combined_num_categ_df[,99] <- NULL  #removing writer column

combined_num_categ_df_train <- combined_num_categ_df 

combined_num_categ_df_train <- combined_num_categ_df_train[, sapply(combined_num_categ_df_train, function(x) is.numeric(x))]

test <- combined_num_categ_df_train[combined_num_categ_df_train$Gross != 0, ]

# Removing linearly dependent or insignificant features
# step has no impact if filterd categorical features are chosen
test$log_month <- NULL
test$square_month <- NULL
test$cube_month <- NULL
test$tomatoImage_rotten <- NULL
test$cube_Released_year <- NULL
test$`Rated_TV-14` <- NULL
test$`Rated_TV-G` <- NULL
test$`Rated_TV-PG` <- NULL

training_size <- seq(0.1, 1, length.out = 20)
training_size_stats_comb2 <- data.frame(matrix(nrow = 20, ncol =3))
colnames(training_size_stats_comb2) <- c("training_size", "train_rmse", "test_rmse")

for(i in 1:length(training_size)){
  #resutls_allFeatures <- RMSE(combined_num_categ_df_train, training_size[i])
  resutls_allFeatures <- RMSE(test, training_size[i])
  training_size_stats_comb2[i,] <- c(training_size[i], resutls_allFeatures[1], resutls_allFeatures[2])
}

lm_new_num <- lm(Gross~., data = test)
summary(lm_new_num)

rmse_task4 <- melt(training_size_stats_comb2[,c('training_size', "train_rmse", "test_rmse")], id = "training_size")
colnames(rmse_task4) <- c("training_size", "variable", "RMSE")

ggplot(rmse_add, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
ggtitle("this plot has not budget Q and runtime bin ")

# this plot has not budget and runtime bin
```

```{r}
rmse_add <- melt(training_size_stats_comb2[,c('training_size', "train_rmse", "test_rmse")], id = "training_size")
colnames(rmse_add) <- c("training_size", "variable", "RMSE")

ggplot(rmse_add, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
ggtitle("this plot has budget Q and runtime bin ")
```


**Q**: Compare the observed RMSE with Tasks 2 & 3.

**A**: Combined model performed slightly better or equivalent to pure numeric model. 
       When Compared the RMSE values combined model performaed better than Task 2 and significantly better than Task 3. 
       When Compared the R-square values combined model performaed better than Task 2 and significantly better than Task 3(0.76 Vs .45)


## 5. Additional features

Now try creating additional features such as interactions (e.g. `is_genre_comedy` x `is_budget_greater_than_3M`) or deeper analysis of complex variables (e.g. text analysis of full-text columns like `Plot`).

```{r}
# TODO: Build & evaluate model 5 (numeric, non-numeric and additional features)

combined_df2 <- cbind(df$Title, combined_df_wanted) # this combined_df_wanted must have budget quantiles
colnames(combined_df2) <- c("Title", colnames(combined_df_wanted))
combined_additional = merge(combined_df2, df_categ, by = "Title")
combined_additional = na.omit(combined_additional)
combined_additional = rename(combined_additional, c("Gross.x" = "Gross"))
combined_additional$Gross.y <- NULL
combined_additional_train <- combined_additional 

combined_additional_train <- combined_additional_train[, sapply(combined_additional_train, function(x) is.numeric(x))]
combined_additional_train <- combined_additional_train[combined_additional_train$Gross != 0, ]

#### Interaction between features extracted from Genre and Production
Print("The idea is high budget movies coming from big production movies in general tend to have high Grossings. Also Genre and Budget category should have some interaction, the idea is ex: Action movies produced with high budget tends to have high Gross")

print("The features used for creating interaction features are one hot encode from original Genre and Budget. Budget was first split using quantiles and then one-hot encoded")

genre_cols <- 43:54
budget_cols <- 26:29
production_cols <- 1:1
all_cols <- colnames(combined_additional_train)

interaction_df <- data.frame(matrix(NA, nrow = nrow(combined_additional_train), ncol = 0))


for(i in genre_cols){
  genre_colname <- all_cols[i]
  genre_colname <- gsub("Genre_", "", genre_colname)
  for(j in budget_cols){
    budget_colname <- all_cols[j]
    budget_colname <- gsub("budget_quant_", "", budget_colname)
    inter_colname <- paste0("Inter_", genre_colname, "_", budget_colname)
    interaction_df[inter_colname] = combined_additional_train[,i] * combined_additional_train[,j]
  }
}

# for(i in genre_cols){
#   prod_colname <- all_cols[i]
#   prod_colname <- gsub("Production_", "", prod_colname)
#   for(j in budget_cols){
#     budget_colname <- all_cols[j]
#     budget_colname <- gsub("budget_quant_", "", budget_colname)
#     inter_colname <- paste0("Inter_", prod_colname, "_", budget_colname)
#     interaction_df[inter_colname] = combined_additional_train[,i] * combined_additional_train[,j]
#   }
# }

### finding the significance of interaction between Genre and Budget by regression p-values
interaction_df$Gross <- combined_additional_train$Gross

lm_new_num <- lm(Gross~., data = interaction_df)
print(summary(lm_new_num))

inter_signif_cols <- c('Inter_Drama_Ultra-Large-Budget',
'Inter_Comedy_Large-Budget',
'Inter_Comedy_Ultra-Large-Budget',
'Inter_Action_Ultra-Large-Budget',
'Inter_Romance_Ultra-Large-Budget',
'Inter_Adventure_Ultra-Large-Budget',
'Inter_Crime_Ultra-Large-Budget',
'Inter_Thriller_Large-Budget',
'Inter_Thriller_Ultra-Large-Budget',
'Inter_Fantasy_Ultra-Large-Budget',
'Inter_Sci-Fi_Ultra-Large-Budget',
'Inter_Family_Ultra-Large-Budget')

print("Below are significant Interaction terms")
print(inter_signif_cols)

interaction_df <- interaction_df[, inter_signif_cols]

combined_additional_train_inter <- cbind(combined_additional_train, interaction_df)

## modelling
combined_additional_train_inter <- combined_additional_train_inter[, sapply(combined_additional_train_inter, function(x) is.numeric(x))]

test1 <- combined_additional_train_inter[combined_additional_train_inter$Gross != 0, ]

lm_new_num <- lm(Gross~., data = test1)
print(summary(lm_new_num))
# Removing linearly dependent features
# step has no impact if filterd categorical features are chosen
test1$`Rated_TV-14` <- NULL
test1$`Rated_TV-G` <- NULL
test1$`Rated_TV-PG` <- NULL
test1$tomatoImage_rotten <- NULL
test1$cube_Released_year <- NULL


training_size <- seq(0.1, 1, length.out = 20)
training_size_stats_add <- data.frame(matrix(nrow = 20, ncol =3))
colnames(training_size_stats_add) <- c("training_size", "train_rmse", "test_rmse")

for(i in 1:length(training_size)){
  #resutls_allFeatures <- RMSE(combined_num_categ_df_train, training_size[i])
  resutls_allFeatures <- RMSE(test1, training_size[i])
  training_size_stats_add[i,] <- c(training_size[i], resutls_allFeatures[1], resutls_allFeatures[2])
}


rmse_add <- melt(training_size_stats_add[,c('training_size', "train_rmse", "test_rmse")], id = "training_size")
colnames(rmse_add) <- c("training_size", "variable", "RMSE")

ggplot(rmse_add, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
  geom_line(linetype = "dashed") + 
  geom_point()
ggtitle("3/5 train Versus test accuracies with training size")


# test_plot_df <- cbind(training_size_stats_add, training_size_stats_comb)
# test_plot_df[,4] <- NULL #making one training_size column NULL
# 
# rmse_add <- melt(test_plot_df[,c('training_size', "train_rmse", "test_rmse", "train_rmse_allFeatures", "test_rmse_allFeatures")], id = "training_size")
# colnames(rmse_add) <- c("training_size", "variable", "RMSE")
# 
# ggplot(rmse_add, aes(x = training_size, y=RMSE, group = variable, color = variable)) + 
#   geom_line(linetype = "dashed") + 
#   geom_point()
# ggtitle("3/5 train Versus test accuracies with training size")
```

**Q**: Explain what new features you designed and why you chose them.

**A**: Interaction terms between one-hot encoded Genre and Budget.
[69]  "Inter_Drama_Ultra-Large-Budget"    
[70] "Inter_Comedy_Large-Budget"          "Inter_Comedy_Ultra-Large-Budget"    "Inter_Action_Ultra-Large-Budget"   
[73] "Inter_Romance_Ultra-Large-Budget"   "Inter_Adventure_Ultra-Large-Budget" "Inter_Crime_Ultra-Large-Budget"    
[76] "Inter_Thriller_Large-Budget"        "Inter_Thriller_Ultra-Large-Budget"  "Inter_Fantasy_Ultra-Large-Budget"  
[79] "Inter_Sci-Fi_Ultra-Large-Budget"    "Inter_Family_Ultra-Large-Budget"   


**Q**: Comment on the final RMSE values you obtained, and what you learned through the course of this project.

**A**:We learned the theory of linear regression and how to apply it. We learned how to use training data to make predictions

